---
title: "Simulation Study - Chapter Two"
author: "Kelly Mistry"
date: "1/28/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(spatstat) ## for calculating nearest neighbor distances
library(ggplot2)
library(reshape2)
library(cowplot)
```

## Simulation Study

### Questions to answer & considerations

-   Think ecologically for this one, more than management

    -   Options:

        -   random distribution and then 2 or 3 with increasing levels of aggregation
        -   range shifts over time would be interesting to try, the question then would be how well does each method capture this

-   Send a fairly detailed proposal to Sarah for comments by mid-February

-   How do I decide how many fish are caught in each haul?

-   I may need to assume that the survey is directly sampling reality, so I don't have to do two steps of sampling (existing population & how it is sampled by the survey)

-   Look into the literature to see what has been done about comparing survey designs, and if there isn't too much published yet than playing with that may be interesting - ask Cecilia if she thinks this is worth pursuing

-   

```{r set up}


## set (x,y) limits for grid size; min = (0,0)
x_max <- 500
y_max <- 500


## number of fish/clusters
n_fish <- 100


```

#### Randomly distributed locations

```{r case_study_1}

set.seed(818)

## assumes nearest neighbor distances ~ Poisson(lambda)

## locations of fish/clusters
rdm_x_loc <- runif(n_fish, 0, x_max)
rdm_y_loc <- runif(n_fish, 0, y_max)

## quick plot of locations
plot(rdm_x_loc, rdm_y_loc, pch = 16, xlim = c(0, x_max), ylim = c(0, y_max))
abline(h = seq(y_max/5, (y_max/5)*4, y_max/5), lty = "dashed")
abline(v = seq(x_max/5, (x_max/5)*4, x_max/5), lty = "dashed")

## calculate nearest neighbor distances
rdm_nnd <- nndist(rdm_x_loc, rdm_y_loc)
# plot distances
hist(rdm_nnd)
```

#### Single disparate cluster

```{r case_study_2a}
set.seed(818)
## assumes nearest neighbor distances ~ NegBin(size, mu)

## locations of fish/clusters as log-normal
agg_x_loc_a <- rnorm(n_fish, 250, 100)
agg_y_loc_a <- rnorm(n_fish, 250, 100)


## quick plot of locations
plot(agg_x_loc_a, agg_y_loc_a, pch = 16, xlim = c(0, x_max), ylim = c(0, y_max))
abline(h = seq(y_max/5, (y_max/5)*4, y_max/5), lty = "dashed")
abline(v = seq(x_max/5, (x_max/5)*4, x_max/5), lty = "dashed")

## calculate nearest neighbor distances
agg_nnd_a <- nndist(agg_x_loc_a, agg_y_loc_a)
# plot distances
hist(agg_nnd_a)
```

#### Single tight cluster

```{r case_study_2b}
set.seed(818)
## assumes nearest neighbor distances ~ NegBin(size, mu)

## locations of fish/clusters as normal
agg_x_loc_b <- rnorm(n_fish, 250, 20)
agg_y_loc_b <- rnorm(n_fish, 250, 20)

## quick plot of locations
plot(agg_x_loc_b, agg_y_loc_b, pch = 16, xlim = c(0, x_max), ylim = c(0, y_max))
abline(h = seq(y_max/5, (y_max/5)*4, y_max/5), lty = "dashed")
abline(v = seq(x_max/5, (x_max/5)*4, x_max/5), lty = "dashed")

## calculate nearest neighbor distances
agg_nnd_b <- nndist(agg_x_loc_b, agg_y_loc_b)
# plot distances
hist(agg_nnd_b)

```

#### Multiple disparate clusters

```{r case_study_2c}
set.seed(818)

## locations of fish/clusters as normal
agg_x_loc_c <- c(rnorm(n_fish/2, 100, 50), rnorm(n_fish/2, 300, 50))
agg_y_loc_c <- c(rnorm(n_fish/2, 100, 50), rnorm(n_fish/2, 300, 50))

## quick plot of locations
plot(agg_x_loc_c, agg_y_loc_c, pch = 16, xlim = c(0, x_max), ylim = c(0, y_max))
abline(h = seq(y_max/5, (y_max/5)*4, y_max/5), lty = "dashed")
abline(v = seq(x_max/5, (x_max/5)*4, x_max/5), lty = "dashed")

## calculate nearest neighbor distances
agg_nnd_c <- nndist(agg_x_loc_c, agg_y_loc_c)
# plot distances
hist(agg_nnd_c)
```

#### Multiple tight clusters

```{r case_study_2d}
set.seed(818)

## locations of fish/clusters as normal
agg_x_loc_d <- c(rnorm(n_fish/4, 100, 20), rnorm(n_fish/4, 200, 20),
                 rnorm(n_fish/4, 300, 20), rnorm(n_fish/4, 400, 20))
agg_y_loc_d <- c(rnorm(n_fish/4, 100, 20), rnorm(n_fish/4, 200, 20),
                 rnorm(n_fish/4, 300, 20), rnorm(n_fish/4, 400, 20))

## quick plot of locations
plot(agg_x_loc_d, agg_y_loc_d, pch = 16, xlim = c(0, x_max), ylim = c(0, y_max))
abline(h = seq(y_max/5, (y_max/5)*4, y_max/5), lty = "dashed")
abline(v = seq(x_max/5, (x_max/5)*4, x_max/5), lty = "dashed")

## calculate nearest neighbor distances
agg_nnd_d <- nndist(agg_x_loc_d, agg_y_loc_d)
# plot distances
hist(agg_nnd_d)
```

```{r plot_scenarios}


par(mfrow=c(3,2))
plot(rdm_x_loc, rdm_y_loc, pch = 16, xlim = c(0, x_max), ylim = c(0, y_max))
abline(v = seq(y_max/4, (y_max/4)*3, y_max/4))
#abline(h = seq(y_max/4, (y_max/4)*3, y_max/4), lty = "dashed")
text(x = c(10, 135, 260, 385), y = rep(500, 4), labels = c("A", "B", "C", "D"), col = "red", font = 2)
plot(agg_x_loc_a, agg_y_loc_a, pch = 16, xlim = c(0, x_max), ylim = c(0, y_max))
# abline(h = seq(y_max/5, (y_max/5)*4, y_max/5), lty = "dashed")
abline(v = seq(y_max/4, (y_max/4)*3, y_max/4))
text(x = c(10, 135, 260, 385), y = rep(500, 4), labels = c("A", "B", "C", "D"), col = "red", font = 2)
plot(agg_x_loc_b, agg_y_loc_b, pch = 16, xlim = c(0, x_max), ylim = c(0, y_max))
# abline(h = seq(y_max/5, (y_max/5)*4, y_max/5), lty = "dashed")
abline(v = seq(y_max/4, (y_max/4)*3, y_max/4))
text(x = c(10, 135, 260, 385), y = rep(500, 4), labels = c("A", "B", "C", "D"), col = "red", font = 2)
plot(agg_x_loc_c, agg_y_loc_c, pch = 16, xlim = c(0, x_max), ylim = c(0, y_max))
# abline(h = seq(y_max/5, (y_max/5)*4, y_max/5), lty = "dashed")
abline(v = seq(y_max/4, (y_max/4)*3, y_max/4))
text(x = c(10, 135, 260, 385), y = rep(500, 4), labels = c("A", "B", "C", "D"), col = "red", font = 2)
plot(agg_x_loc_d, agg_y_loc_d, pch = 16, xlim = c(0, x_max), ylim = c(0, y_max))
# abline(h = seq(y_max/5, (y_max/5)*4, y_max/5), lty = "dashed")
abline(v = seq(y_max/4, (y_max/4)*3, y_max/4))
text(x = c(10, 135, 260, 385), y = rep(500, 4), labels = c("A", "B", "C", "D"), col = "red", font = 2)


par(mfrow=c(3,2))
hist(rdm_nnd)
hist(agg_nnd_a)
hist(agg_nnd_b)
hist(agg_nnd_c)
hist(agg_nnd_d)

```

#### Harvest fish from simulated locations

Below code is doing some weird things at the moment (supposedly catching more fish than exist), so need to figure out what's going on here

```{r harvest}
set.seed(818)

## assumes "catches" occur in 50x50 blocks, with 60% coverage of the space

x_breaks <- sort(sample(seq(0, 450, 50), 8))
y_breaks <- sort(sample(seq(0, 450, 50), 8))

# Plot above to see where catch squares are:
x_start <- x_breaks
x_end <- c(x_breaks+50)
y_start <- y_breaks
y_end <- c(y_breaks+50)
coverage <- as.data.frame(expand.grid(x_breaks, y_breaks))

ggplot(data = coverage) +
  geom_polygon(aes(x = Var1, y = Var2))


x_dims <- length(x_breaks) - 1
y_dims <- length(y_breaks) - 1

## empty matrix for recording catches
catches <- matrix(NA, x_dims, y_dims)

## loop over grid cells
for(i in 1:x_dims) {
  for(j in 1:y_dims) {
    ## fish within x-limits
    xc <- rdm_x_loc > x_breaks[i] & rdm_x_loc <= x_breaks[i + 1]
    ## fish within y-limits
    yc <- rdm_y_loc > y_breaks[j] & rdm_y_loc <= y_breaks[j + 1]
    ## fish within both limits
    catches[i, length(x_breaks) - j] <- sum(xc & yc)
  }
}


## Putting the above process into a function
catch_fun <- function(x_loc, 
                      y_loc, 
                      x_breaks, 
                      y_breaks) {
  # Setting up looping vectors
  x_dims <- length(x_breaks) - 1
  y_dims <- length(y_breaks) - 1
  
  ## empty matrix for recording catches
  catches <- matrix(NA, x_dims, y_dims)
  
  ## loop over grid cells
  for(i in 1:x_dims) {
    for(j in 1:y_dims) {
      ## fish within x-limits
      xc <- x_loc > x_breaks[i] & x_loc <= x_breaks[i + 1]
      ## fish within y-limits
      yc <- y_loc > y_breaks[j] & y_loc <= y_breaks[j + 1]
      ## fish within both limits
      catches[i, length(x_breaks) - j] <- sum(xc & yc)
    }
  }
  return(catches)
}

catches_r <- catch_fun(rdm_x_loc, rdm_y_loc, x_breaks, y_breaks)
catches_a <- catch_fun(agg_x_loc_a, agg_x_loc_a, x_breaks, y_breaks)
catches_b <- catch_fun(agg_x_loc_b, agg_x_loc_b, x_breaks, y_breaks)
catches_c <- catch_fun(agg_x_loc_c, agg_x_loc_c, x_breaks, y_breaks)
catches_d <- catch_fun(agg_x_loc_d, agg_x_loc_d, x_breaks, y_breaks)


## examine catches
t(catches_r)
t(catches_a)
t(catches_b)
t(catches_c)
t(catches_d)

## check total to see if some fish were excluded
sum(catches_r)
sum(catches_a)
sum(catches_b)
sum(catches_c)
sum(catches_d)
```

#### Setting up time series

Initial settings:

-   30 years for the timeseries, so I can sample every other year and get 15 years to put through the models (which is basically the amount of data that I have with POP and northern rockfish)

-   sigma for the white noise version is set to 1.

```{r timeseries_rw}

#### Random walk 

# Number of years
years <- 30

# Sigma for normal error
sigma_wn <- 5

# List of years, with a matrix with x and y coordinates in each year
rdm_timeseries_wn <- list()
year_matrix <- as.data.frame(matrix(NA, nrow = n_fish, ncol = 2))
colnames(year_matrix) <- c("x_loc", "y_loc")
year_matrix[, 1] <- rdm_x_loc
year_matrix[, 2] <- rdm_y_loc
rdm_timeseries_wn[[1]] <- year_matrix

for(year in 2:years) {
  x_year_error <- rnorm(n_fish, 0, sigma_wn)
  y_year_error <- rnorm(n_fish, 0, sigma_wn)
  x_loc <- rdm_timeseries_wn[[year - 1]][, 1] + x_year_error
  y_loc <- rdm_timeseries_wn[[year - 1]][, 2] + y_year_error
  rdm_timeseries_wn[[year]] <- cbind(x_loc, y_loc)
}
  

# Plotting movements of all fish
all_fish_moves_wn <- list()
for(fish in 1:n_fish) {
  fish_mov <- as.data.frame(matrix(NA, nrow = years, ncol = 4))
  colnames(fish_mov) <- c("x_loc", "y_loc", "fish", "year")
  for(year in 1:years) {
  fish_mov[year, 1] <- rdm_timeseries_wn[[year]][fish, 1]
  fish_mov[year, 2] <- rdm_timeseries_wn[[year]][fish, 2]
  fish_mov[year, 3] <- fish
  fish_mov[, 4] <- c(1:years)
  }
  all_fish_moves_wn[[fish]] <- fish_mov
}

# Melting all fish into 1 dataframe to plot against each other
all_fish_melt_wn <- melt(all_fish_moves_wn, id.vars = c("x_loc", "y_loc", "year"))
all_fish_melt_wn <- all_fish_melt_wn[, c(1:3, 5)]
colnames(all_fish_melt_wn)[4] <- "fish_id"

# Plotting all fish, to see how much they move on the whole 
all_fish_wn_plot <- ggplot(all_fish_melt_wn, aes(x = x_loc, y = y_loc, color = as.factor(fish_id))) +
  geom_point() +
  geom_path() +
  theme(legend.position = "none")

# Plotting a few individual fish to see if there is a pattern to their movement between years (there shouldn't be)
sample_fish_wn_plot <- ggplot(all_fish_melt_wn[all_fish_melt_wn$fish_id %in% c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100),]) +
  geom_point(aes(x = x_loc, y = y_loc, color = year)) +
  geom_path(aes(x = x_loc, y = y_loc, color = year)) +
  facet_wrap(vars(fish_id), scales = "free") +
  theme(legend.position = "bottom")

plot_grid(all_fish_wn_plot, sample_fish_wn_plot, labels = c("A", "B"))

```

```{r timeseries_ds}

### Directional shift
# alpha_slope <- c(-0.9, -0.9)
alpha_int <- c(-5, -5) # x and y values; positive means moving north or east, negative means moving south or west
# Do I want this to have some white noise attached, or for it to be set to the same value throughout the timeseries? I think for clarity, it makes sense to initially have it as a set value rather than drawn from a distribution

# Could use the same sigma as in the random walk, but for now set it separately 
sigma_ds <- 5

rdm_timeseries_ds <- list()
year_matrix <- as.data.frame(matrix(NA, nrow = n_fish, ncol = 2))
colnames(year_matrix) <- c("x_loc", "y_loc")
year_matrix[, 1] <- rdm_x_loc
year_matrix[, 2] <- rdm_y_loc
rdm_timeseries_ds[[1]] <- year_matrix

# Version using an alpha as an intercept, added to each year's estimate
for(year in 2:years) {
  x_year_error <- rnorm(n_fish, 0, sigma_ds)
  y_year_error <- rnorm(n_fish, 0, sigma_ds)
  x_loc <- rdm_timeseries_ds[[year - 1]][, 1] + alpha_int[1] + x_year_error
  y_loc <- rdm_timeseries_ds[[year - 1]][, 2] + alpha_int[2] + y_year_error
  rdm_timeseries_ds[[year]] <- cbind(x_loc, y_loc)
}

# # Version using an alpha as a slope for the year's estimate
# for(year in 2:years) {
#   x_year_error <- rnorm(n_fish, 0, sigma_ds)
#   y_year_error <- rnorm(n_fish, 0, sigma_ds)
#   x_loc <- rdm_timeseries_ds[[year - 1]][, 1]*alpha_slope[1] + x_year_error
#   y_loc <- rdm_timeseries_ds[[year - 1]][, 2]*alpha_slope[2] + y_year_error
#   rdm_timeseries_ds[[year]] <- cbind(x_loc, y_loc)
# }
# 
# # The above does a weird thing where it clusters at 0 - probably best to stay with the intercept version

# Plotting movements of all fish
all_fish_moves_ds <- list()
for(fish in 1:n_fish) {
  fish_mov <- as.data.frame(matrix(NA, nrow = years, ncol = 4))
  colnames(fish_mov) <- c("x_loc", "y_loc", "fish", "year")
  for(year in 1:years) {
  fish_mov[year, 1] <- rdm_timeseries_ds[[year]][fish, 1]
  fish_mov[year, 2] <- rdm_timeseries_ds[[year]][fish, 2]
  fish_mov[year, 3] <- fish
  fish_mov[, 4] <- c(1:years)
  }
  all_fish_moves_ds[[fish]] <- fish_mov
}

# Melting all fish into 1 dataframe to plot against each other
all_fish_melt_ds <- melt(all_fish_moves_ds, id.vars = c("x_loc", "y_loc", "year"))
all_fish_melt_ds <- all_fish_melt_ds[, c(1:3, 5)]
colnames(all_fish_melt_ds)[4] <- "fish_id"

all_fish_ds_plot <- ggplot(all_fish_melt_ds, aes(x = x_loc, y = y_loc, color = as.factor(fish_id))) +
  geom_point() +
  geom_path() +
  theme(legend.position = "none")

# Plotting a few individual fish to see if there is a pattern to their movement between years (there shouldn't be)
sample_fish_ds_plot <- ggplot(all_fish_melt_ds[all_fish_melt_ds$fish_id %in% c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100),]) +
  geom_point(aes(x = x_loc, y = y_loc, color = year)) +
  geom_path(aes(x = x_loc, y = y_loc, color = year)) +
  facet_wrap(vars(fish_id), scales = "free") +
  theme(legend.position = "bottom")

## Putting these plots together to show folks:
plot_grid(all_fish_ds_plot, sample_fish_ds_plot, labels = c("A", "B"))


```
